---
title: "Draft - Race Log"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE,
                      cache=TRUE, autodep=TRUE, cache.comments=FALSE)
library(tidyverse)
library(httr)
library(rvest)
library(jsonlite)
library(lubridate)
```

### Overview
- Grab csv of race names and races
- Update all racers before aggregating data
- Get all races from each racer and log them individually in csv files
- Take csv files and aggregate them into a single data file

##### Endpoint used:
https://data.typeracer.com/games?

- playerId=tr:**id** - id = player name
- n = number of races to grab (in this case all)

### Useful Links
- https://robotwealth.com/how-to-wrangle-json-data-in-r-with-jsonlite-purr-and-dplyr/

### Function to Update Racers and Log Race

``` {r}
# updates 
update_racer = function(id)
{
  path = "https://typeracerdata.com/import";
  parameters = list(username = id)
  
  request = GET(url = path, query = parameters)
}

store_races = function(id, races)
{
  # Endpoint will feed in id and total races
  path = "https://data.typeracer.com/games";
  parameters = list(
    playerId = str_c("tr:", id),
    n = 3500 #races (sometimes races aren't updated)
  )
  
  tryCatch(
    {
      request = GET(url = path, query = parameters)
  
      # Convert request json to tibble, some renaming for clarity
      # General data cleanup, removing uneeded columns and converting datetimes
      races = as_tibble(fromJSON(content(request, 'text'))) %>% 
        rename(
          accuracy = ac,
          time = t,
          game_num = gn,
          racers_num = np,
          points = pts
        ) %>% 
        select(everything(), -r, -sl, -tid) %>% 
        mutate(time = as_datetime(time))
      
      # Write to csv in race_log folder with id
      write_csv(races, str_c("race_log/", id, ".csv"))
    },
    error = function(e) {
      # If encounters error with parsing, then sleep for 60 seconds and try again
      print(str_c("error with ", id))
      print(e)
      Sys.sleep(60)
      store_races(id, races)
    }
  )
  
  return ( NULL )
}
```

### Iterate through names and update/log all races
``` {r eval = FALSE}
names = read_csv("names.csv")

for (i in seq_len(nrow(names))) {
  update_racer(names$racer[i])
  print(i)
}

print("all racers logged")

for (i in seq_len(nrow(names))) {
  store_races(names$racer[i], names$races[i])
  print(i)
}
```


### Combine all individual csv files into large csv file
``` {r eval = FALSE} 
combine_racer_data = function(id) {
  data = read_csv(str_c("race_log/", id, ".csv"), show_col_types = FALSE)
  data = data %>% add_column(name = id) %>% select(name, everything())
  data
}

data = tibble()
names = read_csv("names.csv")
for (i in seq_len(nrow(names))) {
  data = data %>% bind_rows(combine_racer_data(names$racer[i]))
}

head(data)

write_csv(data, "all_races.csv")
```
